---
title: "Database Connections"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Database Connections}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Prerequisites

Before using `pifsc.odbc` you need to complete two one-time setup steps:
installing the Oracle ODBC driver and storing your database credentials.

You must also be connected to the **PIFSC network** or the **NOAA VPN** for
all database connections.

### 1. Install Oracle Instant Client

The package requires an Oracle ODBC driver to communicate with the database.
Download and install
[Oracle Instant Client](https://www.oracle.com/database/technologies/instant-client/downloads.html)
(version 23.9 or later) for your operating system. Make sure to include the
ODBC component during installation.
After installation, verify the driver is available:

```{r}
odbc::odbcListDrivers()
```

You should see an entry named `Oracle in instantclient_23_9` (or similar). If
the driver name differs on your machine, you can pass your driver name to
`create_connection()` via the `driver` argument.

### 2. Store database credentials

`pifsc.odbc` uses the
[keyring](https://r-lib.github.io/keyring/) package to securely store your
Oracle username and password in your operating system's credential manager.
This only needs to be done once per machine — credentials persist across R
sessions.

Run the following in an **interactive** R session:

```{r}
library(pifsc.odbc)
setup_credentials()
```

You will be prompted to enter your username and password. These are stored
under the service names `PIFSC_Logbook_user` and `PIFSC_Logbook_pwd` by
default.

If you prefer different service names (e.g., to manage credentials for
multiple databases), you can specify them during setup:

```{r}
setup_credentials(uid_service = "MY_DB_user", pwd_service = "MY_DB_pwd")
```

You will then need to pass the same service names when creating a connection:

```{r}
con <- create_connection(uid_service = "MY_DB_user", pwd_service = "MY_DB_pwd")
```

## Opening a connection

Once the prerequisites are in place, connecting to the logbook database is
straightforward:

```{r}
library(pifsc.odbc)

con <- create_connection()
```

This returns a standard DBI connection object. You can use it with any
DBI-compatible function:

```{r}
# List available tables
DBI::dbListTables(con)

# Run a query
result <- DBI::dbGetQuery(con, "SELECT * FROM llds.SOME_TABLE WHERE ROWNUM <= 10")
```

### Custom connection parameters

If you need to connect to a different database or use non-default settings,
all parameters can be overridden:

```{r}
con <- create_connection(
  host = "other-db.nmfs.local",
  port = 1521,
  sid = "other.service.name",
  driver = "Oracle in instantclient_21_3",
  timeout = 30
)
```

## Closing a connection

Always close your connection when finished. Use `safe_disconnect()` for
error-tolerant cleanup, which is especially useful in scripts and parallel
workers:

```{r}
safe_disconnect(con)
```

Or use the standard DBI approach:

```{r}
DBI::dbDisconnect(con)
```